title: "ATPA - Generalized Additive Models"

CHUNK 1: Load and plot the traffic data
```{r}
library(tidyverse)
setwd("C:/Users/McKay/Documents/Actuary/ATPA/SOAModule_Notes/Module3")
TrafficData <- read_csv("Data/traffic_data.csv") # Read in the simulated traffic data

ggplot(data = TrafficData) +
  geom_point(mapping = aes(x = Hour, y = Traffic, col = Hold_Out))
```

CHUNK 2: Fit and plot linear regression
```{r}
fit1 <- lm(Traffic ~ Hour, data = TrafficData, subset = !Hold_Out)
TrafficData1 <- bind_cols(TrafficData, Predict = predict(fit1, newdata = TrafficData), Order = 1) # Add a new column to the dataset which contains the predictions (makes graphing easier)

# Plot the data and the predictions
ggplot(data = TrafficData1) +
  geom_point(mapping = aes(x = Hour, y = Traffic, col = Hold_Out)) +
  geom_line(mapping = aes(x = Hour, y = Predict))
```

CHUNK 3: Polynomial regression
```{r}
# Fit various polynomial regressions

fit2 <- lm(Traffic ~ poly(Hour, 2), data = TrafficData, subset = !Hold_Out)
fit3 <- lm(Traffic ~ poly(Hour, 3), data = TrafficData, subset = !Hold_Out)
fit5 <- lm(Traffic ~ poly(Hour, 5), data = TrafficData, subset = !Hold_Out)
fit8 <- lm(Traffic ~ poly(Hour, 8), data = TrafficData, subset = !Hold_Out)
fit10 <- lm(Traffic ~ poly(Hour, 10), data = TrafficData, subset = !Hold_Out)


# Prepare the predictions for plotting

TrafficData2 <- bind_cols(TrafficData, Predict = predict(fit2, newdata = TrafficData), Order = 2)
TrafficData3 <- bind_cols(TrafficData, Predict = predict(fit3, newdata = TrafficData), Order = 3)
TrafficData5 <- bind_cols(TrafficData, Predict = predict(fit5, newdata = TrafficData), Order = 5)
TrafficData8 <- bind_cols(TrafficData, Predict = predict(fit8, newdata = TrafficData), Order = 8)
TrafficData10 <- bind_cols(TrafficData, Predict = predict(fit10, newdata = TrafficData), Order = 10)

AllPredTraffic <- bind_rows(TrafficData1, TrafficData2, TrafficData3, TrafficData5, TrafficData8, TrafficData10)

ggplot(data = AllPredTraffic) +
  geom_point(mapping = aes(x = Hour, y = Traffic, col = Hold_Out)) +
  geom_line(mapping = aes(x = Hour, y = Predict)) +
  facet_wrap("Order") +
  coord_cartesian(ylim = c(0, 2500))
```

CHUNK 4: Polynomial regression with log-transformed responses
```{r}
# Fit various polynomial regressions

TrafficData <- TrafficData %>% mutate(logTraffic = log(Traffic))

Lfit1 <- lm(logTraffic ~ Hour, data = TrafficData, subset = !Hold_Out)
Lfit2 <- lm(logTraffic ~ poly(Hour, 2), data = TrafficData, subset = !Hold_Out)
Lfit3 <- lm(logTraffic ~ poly(Hour, 3), data = TrafficData, subset = !Hold_Out)
Lfit5 <- lm(logTraffic ~ poly(Hour, 5), data = TrafficData, subset = !Hold_Out)
Lfit8 <- lm(logTraffic ~ poly(Hour, 8), data = TrafficData, subset = !Hold_Out)
Lfit10 <- lm(logTraffic ~ poly(Hour, 10), data = TrafficData, subset = !Hold_Out)


# Prepare the predictions for plotting

LTrafficData1 <- bind_cols(TrafficData, Predict = exp(predict(Lfit1, newdata = TrafficData)), Order = 1)
LTrafficData2 <- bind_cols(TrafficData, Predict = exp(predict(Lfit2, newdata = TrafficData)), Order = 2)
LTrafficData3 <- bind_cols(TrafficData, Predict = exp(predict(Lfit3, newdata = TrafficData)), Order = 3)
LTrafficData5 <- bind_cols(TrafficData, Predict = exp(predict(Lfit5, newdata = TrafficData)), Order = 5)
LTrafficData8 <- bind_cols(TrafficData, Predict = exp(predict(Lfit8, newdata = TrafficData)), Order = 8)
LTrafficData10 <- bind_cols(TrafficData, Predict = exp(predict(Lfit10, newdata = TrafficData)), Order = 10)

LAllPredTraffic <- bind_rows(LTrafficData1, LTrafficData2, LTrafficData3, LTrafficData5, LTrafficData8, LTrafficData10)

ggplot(data = LAllPredTraffic) +
  geom_point(mapping = aes(x = Hour, y = Traffic, col = Hold_Out)) +
  geom_line(mapping = aes(x = Hour, y = Predict)) +
  facet_wrap("Order") +
  coord_cartesian(ylim = c(0, 2500))
```

CHUNK 5: Generalized additive models
```{r}
library(mgcv)

fitGAM <- gam(Traffic ~ s(Hour), data = TrafficData, subset = !Hold_Out)
TrafficDataGAM <- bind_cols(TrafficData, Predict = predict(fitGAM, newdata = TrafficData))

ggplot(data = TrafficDataGAM) +
  geom_point(mapping = aes(x = Hour, y = Traffic, col = Hold_Out)) +
  geom_line(mapping = aes(x = Hour, y = Predict))


summary(fitGAM)


```

CHUNK 6: Comparing linear models and GAMs
```{r}
summary(lm(Traffic ~ Hour, data = TrafficData, subset = !Hold_Out))
summary(gam(Traffic ~ Hour, data = TrafficData, subset = !Hold_Out))
```

CHUNK 7: Model comparison
```{r}
AIC(fit1, fit2, fit3, fit5, fit8, fit10, fitGAM)
```

CHUNK 8: Checking the GAM
```{r} 
gam.check(fitGAM)
```

CHUNK 9: Reducing the dimension of the smooth
```{r}

fitGAMFew <- gam(Traffic ~ s(Hour, k = 4), data = TrafficData, subset = !Hold_Out)
TrafficDataGAMFew <- bind_cols(TrafficData, Predict = predict(fitGAMFew, newdata = TrafficData))

ggplot(data = TrafficDataGAMFew) +
  geom_point(mapping = aes(x = Hour, y = Traffic, col = Hold_Out)) +
  geom_line(mapping = aes(x = Hour, y = Predict))

summary(fitGAMFew)

gam.check(fitGAMFew)
```

CHUNK 10: Increasing the dimension of the smooth
```{r}

fitGAMMany <- gam(Traffic ~ s(Hour, k = 17), data = TrafficData, subset = !Hold_Out)
TrafficDataGAMMany <- bind_cols(TrafficData, Predict = predict(fitGAMMany, newdata = TrafficData))

ggplot(data = TrafficDataGAMMany) +
  geom_point(mapping = aes(x = Hour, y = Traffic, col = Hold_Out)) +
  geom_line(mapping = aes(x = Hour, y = Predict))

summary(fitGAMMany)

gam.check(fitGAMMany)
```

CHUNK 11: Variable selection example 
```{r}

set.seed(1234)
N <- 10000 # Sample size

# Generate the explanatory variables
x1 <- rnorm(N)
x2 <- rnorm(N)
x3 <- rnorm(N)
x4 <- rnorm(N)
x5 <- rnorm(N)
x6 <- rnorm(N)
x7 <- rnorm(N)
x8 <- rnorm(N)
x9 <- rnorm(N)
x10 <- rnorm(N)

y <- 3 + 2 * x1 + sin(x2) + rnorm(N, 0, 3) # Generate the response variables, only dependent on x1 and sine(x2)

data_mat <- tibble(y, x1, x2, x3, x4, x5, x6, x7, x8, x9, x10)

(GAMFit <- gam(y ~ s(x1) + s(x2) + s(x3) + s(x4) + s(x5) + s(x6) + s(x7) + s(x8) + s(x9) + s(x10), select = TRUE))

plot(GAMFit)

summary(GAMFit)


```

CHUNK 12: Exercise 3.2.1
```{r}
set.seed(1000)

HotelData <- read_csv("Data/hotel_bookings_small.csv")

# 1. 
HotelData <- HotelData %>% mutate(total_stay = stays_in_week_nights + stays_in_weekend_nights)

# 2,
samps <- sample(1:nrow(HotelData), size = 0.8*nrow(HotelData), replace=FALSE)
HotelData_train <- HotelData[samps,]
HotelData_test <- HotelData[-samps,]

# 3. 
mod_train <- gam(adr ~ s(total_stay), data = HotelData_train)


# 4.
plot(mod_train)
mod_train %>% summary

# 5.
mod_train2 <- gam(adr ~ s(total_stay) + lead_time + adults + market_segment, data = HotelData_train)
mod_train3 <- gam(adr ~ total_stay, data = HotelData_train)
mod_train4 <- gam(adr ~ total_stay + lead_time+ adults + market_segment, data = HotelData_train)

mod_train$gcv.ubre.dev
mod_train2$gcv.ubre.dev
mod_train3$gcv.ubre.dev
mod_train4$gcv.ubre.dev

# 5. 
summary(mod_train)
gam.check(mod_train)
summary(mod_train2)
gam.check(mod_train2)

```

CHUNK 13: Exercise 3.2.1: Solution
```{r}

set.seed(1000)

HotelData <- read_csv("Data/hotel_bookings_small.csv") %>%
  mutate(total_stay = stays_in_week_nights + stays_in_weekend_nights) # 1. create total_stay
HotelData <- HotelData %>%
  mutate(set = sample(c("Train", "Test"), nrow(HotelData), prob = c(0.8, 0.2), replace = TRUE)) # 2. Divide the data into train and test

# 3. Fit a simple GAM
HotelGAM <- gam(adr ~ s(total_stay), data = HotelData, subset = (set == "Train")) 
HotelDataGAM <- bind_cols(HotelData, Predict = predict(HotelGAM, newdata = HotelData))

# 4. Plot the predicted values for the test set
ggplot(data = HotelDataGAM[HotelDataGAM$set == "Test", ]) +
  geom_point(mapping = aes(x = total_stay, y = adr)) +
  geom_line(mapping = aes(x = total_stay, y = Predict), col = "blue")
plot(HotelGAM)

# 5. Fit all four models and obtain MSE
HotelLM1 <- lm(adr ~ total_stay, data = HotelData, subset = (set == "Train"))
HotelLM2 <- lm(adr ~ total_stay + lead_time + adults + market_segment, data = HotelData, subset = (set == "Train"))
HotelGAM1 <- gam(adr ~ s(total_stay), data = HotelData, subset = (set == "Train"))
HotelGAM2 <- gam(adr ~ s(total_stay) + lead_time + adults + market_segment, data = HotelData, subset = (set == "Train"))


HotelAllFit <- bind_cols(HotelData,
  PredLM1 = predict(HotelLM1, newdata = HotelData),
  PredLM2 = predict(HotelLM2, newdata = HotelData),
  PredGAM1 = predict(HotelGAM1, newdata = HotelData),
  PredGAM2 = predict(HotelGAM2, newdata = HotelData)
)

HotelAllFit %>%
  filter(set == "Test") %>%
  mutate(
    errorLM1 = PredLM1 - adr,
    errorLM2 = PredLM2 - adr,
    errorGAM1 = PredGAM1 - adr,
    errorGAM2 = PredGAM2 - adr,
  ) %>%
  summarize(
    MSPELM1 = mean(errorLM1^2),
    MSPELM2 = mean(errorLM2^2),
    MSPEGAM1 = mean(errorGAM1^2),
    MSPEGAM2 = mean(errorGAM2^2)
  )

# 6. Model Diagnostics
gam.check(HotelGAM1)
gam.check(HotelGAM2)

# 7. Analyze the chosen model
summary(HotelGAM2)
```

CHUNK 14 - Gamma GLM on exercise

```{r}

HotelDataGLM <- HotelData %>% 
                  mutate(adr1 = adr + 1)

HotelLM <- lm(adr1 ~ total_stay + lead_time + adults + market_segment, data = HotelDataGLM, subset = (set == "Train"))
HotelLMGAM <- gam(adr1 ~ s(total_stay) + lead_time + adults + market_segment, data = HotelDataGLM, subset = (set == "Train"))
HotelGLM <- glm(adr1 ~ total_stay + lead_time + adults + market_segment, family="Gamma", data = HotelDataGLM, subset = (set == "Train"))
HotelGLMGAM <- gam(adr1 ~ s(total_stay) + lead_time + adults + market_segment, family="Gamma", data = HotelDataGLM, subset = (set == "Train"))

HotelGLMAllFit <- bind_cols(HotelDataGLM,
  PredLM = predict(HotelLM, newdata = HotelDataGLM),
  PredLMGAM = predict(HotelLMGAM, newdata = HotelDataGLM),
  PredGLM = predict(HotelGLM, newdata = HotelDataGLM, type = "response"),
  PredGLMGAM = predict(HotelGLMGAM, newdata = HotelDataGLM, type = "response")
)

HotelGLMAllFit %>%
  filter(set == "Test") %>%
  mutate(
    errorLM = PredLM - adr1,
    errorLMGAM = PredLMGAM - adr1,
    errorGLM = PredGLM - adr1,
    errorGLMGAM = PredGLMGAM - adr1,
  ) %>%
  summarize(
    MSPELM = mean(errorLM^2),
    MSPELMGAM = mean(errorLMGAM^2),
    MSPEGLM = mean(errorGLM^2),
    MSPEGLMGAM = mean(errorGLMGAM^2)
  )

```
