---
title: "ATPA Linear Mixed Models"
---

CHUNK 1: Read in and plot data
```{r}
library(tidyverse)
comp_course <- read_csv("Data/comp_course.csv")
comp_course %>% ggplot(aes(x = grade)) +
  geom_histogram(color = "darkblue", fill = "lightblue")
```

CHUNK 2: Fit standard regression model
```{r}
lm_mod <- lm(grade ~ ., data = comp_course)
summary(lm_mod)
drop1(lm_mod) # Uses AIC to determine if variables can be dropped, treating categorical variables as a single variable to keep or drop. All three predictors are deemed significant.
```

CHUNK 3: Fit a random intercepts model 
```{r}
# the lme4 package is used for fitting mixed models
# install.packages("lme4")
library(lme4)
# A variable is treated as a random effect by using (1|variable) in the model formula
ri_mod <- lmer(grade ~ (1 | prof) + year + hours + major, data = comp_course)
```

CHUNK 4: Compare fits and predictions between standard and mixed model
```{r}
# Compare coefficients between standard and mixed model
cbind(lm_mod$coefficients[6:10], summary(ri_mod)$coefficients[-1, 1])

# Find and plot the predictions for the two models.
plt_dat <- tibble("lm_preds" = predict(lm_mod), "mm_preds" = predict(ri_mod))
plt_dat %>% ggplot(aes(x = lm_preds, y = mm_preds)) +
  geom_point()
```

CHUNK 5: Find random effects
```{r}
# use the ranef function to get random effects of a mixed model
ranef(ri_mod)

# Linear Regression model Effects
lm_prof <- c(0, lm_mod$coefficients[2:5])
prof_effects <- lm_prof - mean(lm_prof)
prof_effects
```

CHUNK 6: Perform prediction for new data with a new factor level for Professor
```{r}
comp_course_2 <- read_csv("Data/comp_course_2.csv")

# We need to split this data into two parts, the observations with the same professors as the original data and the observations with new professors
oldprof <- comp_course_2$prof %in% unique(comp_course$prof)
# Predict all data using argument "re.form=NA" which says to not add mean for random effect. This will be best for observations with new factor levels
preds <- predict(ri_mod, newdata = comp_course_2, re.form = NA)
# Replace observations with professors from original data with predictions having added in the mean of the random effect (because it exists only for these ones)
preds[oldprof] <- predict(ri_mod, newdata = comp_course_2[oldprof, ])

# Plot the predictions
isF <- ifelse(comp_course_2$prof == "F", "F", "not F")
tibble("Grade" = comp_course_2$grade, "Prediction" = preds, "Professor" = isF) %>%
  ggplot(aes(x = Grade, y = Prediction)) +
  geom_point(aes(color = Professor))
```

CHUNK 7: Fit a random slopes model 
```{r}

# The new interaction term is "(0+fixed_effect|random_effect)". In this case it is "(0+hours|prof)". hours is still included as a fixed effects variable
rs_mod <- lmer(grade ~ (1 | prof) + (0 + hours | prof) + year + hours + major, data = comp_course)

# The summary has several parts to it. The most important pieces are (1) the random effects part gives the variance of each random effect and (2) the fixed effects part gives estimates and standard errors for the fixed effect variables
summary(rs_mod)

# Notice that hours shows up as a fixed effect but also as an interaction with prof as a random effect
```

CHUNK 8: Create data for Exercise 3.3.2
```{r}
individual <- rep(c("Aaron", "Felipe", "Sonia"), each = 5)
weekend <- c(1, 1, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 1, 1)
weather <- c("sunny", "cloudy", "rain", "other", "sunny", "sunny", "rain", "rain", "cloudy", "other", "sunny", "sunny", "sunny", "rain", "cloudy")
temperature <- c(79, 60, 72, 76, 69, 92, 80, 72, 79, 81, 56, 61, 68, 78, 79)
email_sent <- c(5, 6, 14, 17, 24, 8, 0, 1, 5, 12, 25, 22, 21, 12, 13)
email <- tibble(individual, weekend, weather, temperature, email_sent)
email

# Part 1
"Individual - Random
 Others - Fixed"

# Part 2
ranint_mod <- lmer(email_sent ~ (1 | individual) + weekend + weather + temperature, data = email)

# Part 3
ranintslopes_mod <- lmer(email_sent ~ (1 | individual) + (0 + temperature | individual) + weekend + weather + temperature, data = email)
```

CHUNK 9: Exercise 3.3.2: Solution
```{r}
individual <- rep(c("Aaron", "Felipe", "Sonia"), each = 5)
weekend <- c(1, 1, 0, 0, 0, 0, 1, 1, 0, 0, 0, 0, 0, 1, 1)
weather <- c("sunny", "cloudy", "rain", "other", "sunny", "sunny", "rain", "rain", "cloudy", "other", "sunny", "sunny", "sunny", "rain", "cloudy")
temperature <- c(79, 60, 72, 76, 69, 92, 80, 72, 79, 81, 56, 61, 68, 78, 79)
email_sent <- c(5, 6, 14, 17, 24, 8, 0, 1, 5, 12, 25, 22, 21, 12, 13)
email <- tibble(individual, weekend, weather, temperature, email_sent)
email

# Part 2
email_ri <- lmer(email_sent ~ (1 | individual) + weekend + weather + temperature, data = email)
summary(email_ri)

# Part 3
email_rs <- lmer(email_sent ~ (1 | individual) + (0 + temperature | individual) + weekend + weather + temperature, data = email)
summary(email_rs)
```

CHUNK 10: Create data frame and fit repeated measures and longitudinal models
```{r}
Patient <- rep(c(1, 2, 3), each = 4)
Time <- rep(c(0, 10, 20, 30), 3)
Dosage <- c(0, 60, 90, 100, 0, 30, 50, 70, 0, 40, 80, 100)
Response <- c(710, 540, 510, 450, 650, 590, 500, 460, 890, 700, 590, 440)
dose_response <- tibble(Patient, Time, Dosage, Response)

# Fit random slopes model without time
dose_mod1 <- lmer(Response ~ (1 | Patient) + Dosage, data = dose_response)

# Add time as a fixed effect
dose_mod2 <- lmer(Response ~ (1 | Patient) + Dosage + Time, data = dose_response)

# Create a longitudinal model by adding an interaction between patient and time
dose_mod3 <- lmer(Response ~ (1 | Patient) + (0 + Time | Patient) + Dosage + Time, data = dose_response)

# Compare the model fit using AIC function
AIC(dose_mod1)
AIC(dose_mod2)
AIC(dose_mod3)

# The model with an interaction between time and the random effect is the best
```

CHUNK 11: Revisit computer course data with a GLMM
```{r}
comp_course$got_A <- comp_course$grade >= 90

# Just like in the glm function, a new argument for family is included with the same options
glmm_mod_ri <- glmer(got_A ~ (1 | prof) + year + hours + major, data = comp_course, family = binomial(link = "logit"))


# Random slopes
glmm_mod_rs <- glmer(got_A ~ (1 | prof) + (0 + hours | prof) + year + hours + major, data = comp_course, family = binomial(link = "logit"))

# Fixed effect only
glm_mod <- glm(got_A ~ prof + year + hours + major + prof:hours, data = comp_course, family = binomial(link = "logit"))

## Compare all three models using AIC
AIC(glmm_mod_ri)
AIC(glmm_mod_rs)
AIC(glm_mod)

# Fixed effect model is actually the best fit for this model. Could use other metrics, such as prediction for test or validation sets to confirm. There are other reasons to use mixed models beyond model fit, as mentioned.
```

CHUNK 12: Exercise 3.3.3
```{r}

individual <- rep(c("Aaron", "Felipe", "Sonia"), each = 5)
day <- rep(1:5, 3)
weather <- c("sunny", "cloudy", "rain", "other", "sunny", "sunny", "rain", "rain", "cloudy", "other", "sunny", "sunny", "sunny", "rain", "cloudy")
temperature <- c(79, 60, 72, 76, 69, 92, 80, 72, 79, 81, 56, 61, 68, 78, 79)
email_sent <- c(5, 6, 14, 17, 24, 8, 0, 1, 5, 12, 25, 22, 21, 12, 13)
email <- tibble(individual, day, weather, temperature, email_sent)
# email

mod1 <- lmer(email_sent ~ (1 | individual) + (0 + day | individual) + day + weather + temperature, data = email)

mod_glmm <- glmer(email_sent ~ (1 | individual) + (0 + day | individual) + day + weather + temperature, data = email, family = poisson(link = 'log'))

AIC(mod1)
AIC(mod_glmm)

```

CHUNK 13: Exercise 3.3.3: Solution
```{r}
individual <- rep(c("Aaron", "Felipe", "Sonia"), each = 5)
day <- rep(1:5, 3)
weather <- c("sunny", "cloudy", "rain", "other", "sunny", "sunny", "rain", "rain", "cloudy", "other", "sunny", "sunny", "sunny", "rain", "cloudy")
temperature <- c(79, 60, 72, 76, 69, 92, 80, 72, 79, 81, 56, 61, 68, 78, 79)
email_sent <- c(5, 6, 14, 17, 24, 8, 0, 1, 5, 12, 25, 22, 21, 12, 13)
email <- tibble(individual, day, weather, temperature, email_sent)

# random slopes model for the email data
email_rs2 <- lmer(email_sent ~ (1 | individual) + (0 + day | individual) + day + weather + temperature, data = email)

# glmm random slopes model
email_rs3 <- glmer(email_sent ~ (1 | individual) + (0 + day | individual) + day + weather + temperature, data = email, family = poisson(link = "log"))

AIC(email_rs2)
AIC(email_rs3)

# The normal model is the better fit for this data
```

CHUNK 14: Create the loss data set and calculate the credibility premium
```{r}
line <- c(1, 1, 1, 2, 2, 2, 3, 3, 3)
exposure <- c(20, 30, 40, 10, 5, 10, 100, 70, 75)
loss <- c(4000, 4500, 7000, 2000, 1500, 2500, 15000, 13000, 12000)

# calculate average loss per exposure
avloss <- loss / exposure

# build data for building a model
cred_dat <- tibble(line, exposure, avloss)

# build a mixed model and view random effects
mod_cred <- lmer(avloss ~ 1 + (1 | line), data = cred_dat, weights = exposure)
mod_cred
ranef(mod_cred)

# examine predictions for each line of business
cred_premium <- unique(cbind(line, predict(mod_cred)))
cred_premium

# Total predicted claims
sum(cred_dat$exposure*predict(mod_cred))
loss %>% sum
```

CHUNK 15: Calculate credibility premium with gamma distribution
```{r}
mod_cred2 <- glmer(avloss ~ 1 + (1 | line), data = cred_dat, weights = exposure, family = Gamma(link = "log"))
cred_premium2 <- unique(cbind(line, predict(mod_cred2, type = "response")))
cred_premium2
```

CHUNK 16: Exercise 3.3.4
```{r}
claims <- c(12, 12, 20, 8, 2, 7, 20, 21, 13)
cred_dat$avclaims <- claims / cred_dat$exposure

mod <- lmer(avclaims ~ 1 + (1 | line), data = cred_dat, weights = exposure)
unique(bind_cols(Line = cred_dat$line, Preds = predict(mod)))
```

CHUNK 17: Exercise 3.3.4: Solution
```{r}
claims <- c(12, 12, 20, 8, 2, 7, 20, 21, 13)
cred_dat$avclaims <- claims / cred_dat$exposure

mod_cred3 <- lmer(avclaims ~ 1 + (1 | line), data = cred_dat, weights = exposure)
cred_premium3 <- unique(cbind(line, predict(mod_cred3)))
cred_premium3
```
